### 2025-06-28 manuscript/main.tex
part division - highlevel architecture
begin explaination of components

\usepackage{forest}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[colorlinks,linkcolor=blue,urlcolor=cyan]{hyperref}
\usepackage{titlesec}
\titleformat{\chapter}[hang]{\normalfont\huge\bfseries}{}{0pt}{}
%=== Metadata ===
\newcommand{\thesistitle}{Your Thesis Title Here}
\newcommand{\authorname}{Rafael Tanzer}
\newcommand{\supervisor}{Michele Chiaria, Simone , Francesco Pontigia and Ezzio Bartocci}
\newcommand{\institution}{TU Vienna}
\newcommand{\department}{Cyber-Physical-Systems, ..., ...}
\newcommand{\logoimage}{graphics/logo.png} % Path to your logo image
\newcommand{\submissionmonth}{May}
\newcommand{\submissionyear}{2025}
%================

@@

%=== References ===%
\cleardoublepage
\part{References}
\printbibliography[heading=subbibliography,title={References for Part I},section=1]
\printbibliography[heading=subbibliography,title={References for Part II},section=2]

@@

\chapter{Test Data}
% Sample inputs, test cases, etc.

>>>>

\usepackage{forest}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[colorlinks,linkcolor=blue,urlcolor=cyan]{hyperref}
\usepackage{titlesec}
\titleformat{\chapter}[hang]{\normalfont\huge\bfseries}{}{0pt}{}
\titlespacing*{\chapter}{0pt}{1.5em}{1em}
%=== Metadata ===
\newcommand{\thesistitle}{Your Thesis Title Here}
\newcommand{\authorname}{Rafael Tanzer}
\newcommand{\supervisor}{Michele Chiaria, Simone , Francesco Pontigia and Ezzio Bartocci}
\newcommand{\institution}{TU Vienna}
\newcommand{\department}{Cyber-Physical-Systems, ..., ...}
\newcommand{\logoimage}{graphics/logo.png} % Path to your logo image
\newcommand{\submissionmonth}{May}
\newcommand{\submissionyear}{2025}
%================

@@

%=== References ===%
\cleardoublepage
\printbibliography[heading=subbibliography,title={Theoretical Foundation},section=1]
\printbibliography[heading=subbibliography,title={Technological Implementation},section=2]

@@

### 2025-06-28 manuscript/part2.tex
notes for travel
begin explaination of components

%=== Part II - Design \& Implementation ===%
\part{Design \& Implementation}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{graphics/langiumArchitecture.png}
  \caption{Architecture overview of a Langium project, showing core services, entry points, and integrations.}
  \label{fig:langium-architecture}
\end{figure}

@@

@@

>>>>

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{graphics/langiumArchitecture.png}
  \caption{Architecture overview of a Langium project}
  \label{fig:langium-architecture}
\end{figure}

@@

Langium projects can easily be created with the support of the scaffolding tool \verb|Yeoman|, producing a structured mock project including a sample language definition.
Most of the definitions are found in the \verb|src| directory \ref{fig:langium-architecture}, encompassing fnctionality for the Language Server, the extension and actual language tooling.

\begin{description}
  \item[\code{language/}] Assembles the core of any Langium project. It includes the \code{.langium} file, in which the custom language grammar is defined, as well as the code that starts the Language Server with all necessary services bundled.
  \item[\code{cli/}] Provides functionality for CLI interactions, command parsing, and generators. The generators in this folder invoke processes that generate artifacts from the language and input files;
  they are typically used only offline and are called once rather than continuously via the CLI and support efficient pipelining.
  \item[\code{extension/}] Contains the VS Code extension bootstrap code needed to connect to the Language Server and utilize LSP support features. All folders except the CLI/ are bundled into the extension itself.
  \item[\code{generated/}] Contains artifacts produced from the grammar definition in the \code{.langium} file. This includes TypeScript AST type definitions, which are essential for further analysis and validation, as well as the preliminary parsed grammar rules formatted for the Chevrotain engine to generate parsers.
  \item[\textbf{Services}] The Services in Langium are a collection of utility classes and helper functions offered to every component of a language project via dependency injection. They cover an extensive range of functionality, from LSP-specific features - such as providers for context-aware code completion, rich hover tooltips, signature help, go-to-definition and find-references - to core language utilities that expose parsed documents as both AST and CST models, manage workspace state and context information (including indexing, scoping, and name resolution), synchronize the in-memory document model and perform validation checks. Because the parts are wired up as an injectable service, precise consumptions of the capabilities needed is possible. Morover it Langium enables developers to swap in custom implementations to extend the language's behavior. ..expand on CustpmServices
\end{description}

Modules  .... register services, expose interfaces

Although the crucial segments are located inside \code{/src}, files of interest ... Szntax config

@@

Validation.ts  was already placed in /language, subsequenntly, other custom services also implementd here

Lastly, situated at top-level in the \verb|src| directory are three files: \verb|setup....ts| \verb|setup....ts| \verb|setup....ts|, which are invoked
by
